---
title: "extract_xy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{extract_xy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```


```{r setup}
library(xyLIDAR)
library(tidyverse)

# library(BIOMASS)
# library(plotsdatabase)
# library(data.table)
# library(grDevices)
# library(sf)
# library("xlsx")
```

Set the parameters 

```{r}
# Parameters ####

root_in = system.file(package="xyLIDAR", 'mbalmayo001') #path to the main folder of the lidar scans files
plot_name = "mbalmayo001" # plotsdatabase format
method = 'plotsdatabase' # Choose 'plotdatabase' or 'local_file'
local_file = NULL # path to the local file inventory data
crs = 'EPSG:32633'
directory <- 'E:/VIA/scan_IPhone/outputs'

dir.create(file.path(directory,plot_name))
load(system.file(package="xyLIDAR", 'color_plot.RData') )
```

Import all the the lidar data files within a list (list_of_Lidar). Each element correspond to one file with 3 variables (id, X_lidar, Y_lidar)

files_summary resumes the informations included in each files. If the infomation 
are not well organized in the dataframe, it means that the file is not well named.

```{r}
summary <- files_summary(root_in,
                         export = TRUE, 
                         directory = directory, 
                         plot_name = plot_name)

list_of_Lidar <- summary$list_of_Lidar
files_summary <- summary$file_summary

```


Extract field data 

```{r}
Fieldplot_BDD_full <- import_plotdata(method,plot_name) # password : Amap2020, username : ploton
```

Compute jalons coordinates 

```{r}
subplot <- extract_jalons(coordinates = Fieldplot_BDD_full$coordinates)

library(sf)
test3 = subplot %>% 
   st_as_sf(coords = c('XAbs', 'YAbs')) 
test3 <- test3 
library(mapview)
mapview(test3)%>% 
   leafem::addStaticLabels(label = test3$jalon,
                           noHide = TRUE,
                           direction = 'top',
                           textOnly = TRUE,
                           textsize = "20px")




subplot_recode <- function(subplot) {
   subplot$jalon <- forcats::fct_recode(subplot$jalon,
                                  "0_0" = '100_0',
                                  "20_0" = '80_0',
                                  "40_0" = '60_0',
                                  "60_0" = '40_0',
                                  "80_0" = '20_0',
                                  "100_0" = '0_0',
                                  "0_20" = '100_20',
                                  "20_20" = '80_20',
                                  "40_20" = '60_20',
                                  "60_20" = '40_20',
                                  "80_20" = '20_20',
                                  "100_20" = '0_20',
                                  "0_40" = '100_40',
                                  "20_40" = '80_40',
                                  "40_40" = '60_40',
                                  "60_40" = '40_40',
                                  "80_40" = '20_40',
                                  "100_40" = '0_40',
                                  "0_60" = '100_60',
                                  "20_60" = '80_60',
                                  "40_60" = '60_60',
                                  "60_60" = '40_60',
                                  "80_60" = '20_60',
                                  "100_60" = '0_60',
                                  "0_80" = '100_80',
                                  "20_80" = '80_80',
                                  "40_80" = '60_80',
                                  "60_80" = '40_80',
                                  "80_80" = '20_80',
                                  "100_80" = '0_80',
                                  "0_100" = '100_100',
                                  "20_100" = '80_100',
                                  "40_100" = '60_100',
                                  "60_100" = '40_100',
                                  "80_100" = '20_100',
                                  "100_100" = '0_100'
   )
   

   
   subplot$sousplot <- forcats::fct_recode(subplot$sousplot,
                                  "80_0" = '0_0',
                                  "60_0" = '20_0',
                                  "40_0" = '40_0',
                                  "20_0" = '60_0',
                                  "0_0" = '80_0',
                                  "80_20" = '0_20',
                                  "60_20" = '20_20',
                                  "40_20" = '40_20',
                                  "20_20" = '60_20',
                                  "0_20" = '80_20',
                                  "80_40" = '0_40',
                                  "60_40" = '20_40',
                                  "40_40" = '40_40',
                                  "20_40" = '60_40',
                                  "0_40" = '80_40',
                                  "80_60" = '0_60',
                                  "60_60" = '20_60',
                                  "40_60" = '40_60',
                                  "20_60" = '60_60',
                                  "0_60" = '80_60',
                                  "80_80" = '0_80',
                                  "60_80" = '20_80',
                                  "40_80" = '40_80',
                                  "20_80" = '60_80',
                                  "0_80" = '80_80',
   )
      subplot$XRel <- as.numeric(str_split(subplot$jalon,'_',simplify = TRUE)[,1])
   return(subplot)
}

subplot <- subplot_recode(subplot)


```

Reproject trees using the jalons coordinates as reference 

```{r}
files_summary_jalon <- 
   files_summary %>% 
   filter(nb_jalon > 2)

dir.create(file.path(directory,plot_name,"xy_rawdata_method1"))
dir.create(file.path(directory,plot_name,"xy_reproject_method1"))

for(i in 1:nrow(files_summary_jalon) ){
   
   file_info <- files_summary_jalon[i,]
   
   LIDAR_sousplot <- 
      XY_computation(
      Fieldplot_BDD_full,
      list_of_Lidar,
      subplot,
      plot_name,
      file_info,
      method = 'jalon',
      directory)
      
      
      assign(  paste("Reproj_XY_jalon", unique(LIDAR_sousplot$file), sep = "_"), LIDAR_sousplot )

}

list_of_objects <- mget(ls(pattern="Reproj_XY_jalon"))
all_jalon_method <- do.call("rbind", list_of_objects)
```

Reproject trees using the jalons and the reprojected trees coordinates as reference 

```{r}
files_summary_trees <- 
   files_summary %>% 
   filter(nb_jalon < 3)

dir.create(file.path(directory,plot_name,"xy_rawdata_method2"))
dir.create(file.path(directory,plot_name,"xy_reproject_method2"))

for(i in 1:nrow(files_summary_trees) ){
   
   file_info <- files_summary_trees[i,]
   
   LIDAR_sousplot <- 
      XY_computation(
      Fieldplot_BDD_full,
      list_of_Lidar,
      subplot,
      plot_name,
      file_info,
      method = 'tree',
      directory)
      
      
      assign(  paste("Reproj_XY_tree", unique(LIDAR_sousplot$file), sep = "_"), LIDAR_sousplot )

}

list_of_objects <- mget(ls(pattern="Reproj_XY_tree"))
all_tree_method <- do.call("rbind", list_of_objects)

```

```{r}
all <- rbind(all_jalon_method, all_tree_method)
```

Plot all trees

```{r}

sub_plot <- plot_alltrees(
   Fieldplot_BDD_full,
   all,
   subplot,
   crs,
   col)

```


```{r}
save_gpkg(all, crs, sub_plot)
```

```{r}
{
   sink(file = paste0(plot_name,"/5_ALL_PLOT_REPORT.txt"))

   all_missing_id <- unique(Fieldplot_BDD_full$extract$ind_num_sous_plot)[!(unique(Fieldplot_BDD_full$extract$ind_num_sous_plot) %in% unique(all$id))] %>% as.numeric %>% sort()
   aa <- all %>% filter(where == 'in') %>% .[['id']] %>% unique()
   all_missing_id_in <- unique(Fieldplot_BDD_full$extract$ind_num_sous_plot)[!(unique(Fieldplot_BDD_full$extract$ind_num_sous_plot) %in% aa)] %>% as.numeric %>% sort()


   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('                       MISSING ID IN THE SCAN                       '))
   print(paste('####################################################################'))
   print(paste('                                                                    '))


   print(paste('The following id are not present on the scans :', paste(all_missing_id_in, collapse = ', ')))


   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('    MISSING ID IN THE SCAN AND THE SUBPLOT TO BE (RE)SCANNED        '))
   print(paste('####################################################################'))
   print(paste('                                                                    '))


   print(Fieldplot_BDD_full$extract %>% filter(ind_num_sous_plot %in% all_missing_id_in) %>% group_by(sous_plot_name) %>%  summarise(id = paste(ind_num_sous_plot, collapse = ', ')))


   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('        MISSING ID CAN BE DUE TO THE FOLLOWING MISTAKE              '))
   print(paste('####################################################################'))
   print(paste('                                                                    '))

   adjacent_error <- all %>% filter(where == 'adjacent')
   out_error <- all %>% filter(where == 'out')

   if(nrow(adjacent_error) > 0) {

      for (i in 1:nrow(adjacent_error)) {

         print(paste('In the file :', adjacent_error$file[i], ', the id', adjacent_error$id[i], ' is present but it is suppose to be on the adjacent subplot', adjacent_error$sous_plot_name[i]))

      }
   }

   if(nrow(out_error) > 0) {

      for (i in 1:nrow(out_error)) {

         print(paste('In the file :', out_error$file[i], ', the id', out_error$id[i], ' is present but it is suppose to be on the another subplot which is not adjacent', out_error$sous_plot_name[i]))

      }
   }


   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('####################################################################'))
   print(paste('                                                                    '))

   duplicate_id_same_file <- all %>%
      filter(duplicated_id == 'yes' & what == 'tree') %>%
      group_by(id) %>%
      summarise(file = unique(file),
                n = n())

   if (nrow(duplicate_id_same_file) > 0) {


      for (i in 1:nrow(duplicate_id_same_file)){
         print(paste('In the file :', duplicate_id_same_file$file[i], ', the id', duplicate_id_same_file$id[i], ' is present', duplicate_id_same_file$n[i], ' times'))

      }


   }

   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('                FILES WITH LESS THAN 3 REFERENCES                   '))
   print(paste('####################################################################'))
   print(paste('                                                                    '))

   less_than_3_ref <- all %>%
      group_by(file) %>%
      slice(1) %>%
      select(file, n_jalon_ref, n_tree_ref, n_tot_ref) %>%
      filter(n_tot_ref < 3)

   if (nrow(less_than_3_ref) > 0) {


      for (i in 1:nrow(less_than_3_ref)){
         print(paste('The file :', less_than_3_ref$file[i], 'has only', less_than_3_ref$n_tot_ref[i], ' reference(s)'))
         print(paste('It has :', less_than_3_ref$n_jalon_ref[i], ' jalon(s)'))
         print(paste('It has :', less_than_3_ref$n_tree_ref[i], ' tree(s)'))
         print(paste('                                                                    '))

      }


   }

   print(paste('                                                                    '))
   print(paste('####################################################################'))
   print(paste('                  FILES WITH LESS THAN 3 JALONS                     '))
   print(paste('####################################################################'))
   print(paste('                                                                    '))

   less_than_3_jalon <- all %>%
      group_by(file) %>%
      slice(1) %>%
      select(file, n_jalon_ref, n_tree_ref, n_tot_ref) %>%
      filter(n_jalon_ref < 3)

   if (nrow(less_than_3_jalon) > 0) {


      for (i in 1:nrow(less_than_3_jalon)){
         print(paste('The file :', less_than_3_jalon$file[i], 'has only', less_than_3_jalon$n_jalon_ref[i], ' jalon(s) as reference(s)'))
         print(paste('It has :', less_than_3_jalon$n_jalon_ref[i], ' jalon(s)'))
         print(paste('It has :', less_than_3_jalon$n_tree_ref[i], ' tree(s)'))
         print(paste('                                                                    '))

      }


   }

   sink(file =NULL)

   print(paste('A report has been saved : ', paste0(getwd(), '/',plot_name,"/5_ALL_PLOT_REPORT.txt")))

} # ALL SCANS REPORT
```

```{r}
dir.create(paste0(directory,'/',plot_name,"/20_20_PROJECTION"))

{

   for (i in 1:length(unique(all$file))) {

      tmp <- all %>% filter(file == unique(all$file)[i])
      jalon_ref <- subplot %>% filter(sousplot == unique(tmp$sousplot))
      jalon_ref <- jalon_ref[as.character(jalon_ref$jalon) %in% tmp$id,]

      jalon_ref <-jalon_ref %>% left_join(tmp %>% filter(what == 'jalon' & where == 'in' ) %>% select(-c("XAbs", "YAbs")), by = join_by(jalon == id))

      res <-procrust((jalon_ref[,c("XRel", "YRel")]), jalon_ref[,c("XAbs", "YAbs")])


      coordAbs_allTrees <- as.matrix(tmp[,c("XAbs", "YAbs")]) %*% res$rotation
      coordAbs_allTrees <- sweep(coordAbs_allTrees, 2, res$translation, FUN = "+")
      tmp$TRUE_X <- coordAbs_allTrees[,1]
      tmp$TRUE_Y <- coordAbs_allTrees[,2]

      TRUE_XY <- tmp
      assign(  paste("TRUE_XY_20_20", unique(TRUE_XY$file), sep = "_"), TRUE_XY )

   }

   list_of_objects <- mget(ls(pattern="TRUE_XY_20_20"))

   # bind the elements together into a data frame

   all <- do.call("rbind", list_of_objects)

   all$quadrat_x <- as.numeric(str_split(all$sousplot, '_', simplify = TRUE)[,1])
   all$quadrat_y <- as.numeric(str_split(all$sousplot, '_', simplify = TRUE)[,2])

   all <- all %>%
      mutate(TRUE_X = TRUE_X - quadrat_x,
             TRUE_Y = TRUE_Y - quadrat_y) %>% print(n=100)

   for (i in 1:length(unique(all$sousplot))){


      tmp <- all %>% filter(sousplot == unique(all$sousplot)[i])

      my_plot <- ggplot(tmp) +
         ggrepel::geom_label_repel(aes(label = id,x=TRUE_X, y=TRUE_Y, col = where, fill = what), size = 5) +
         scale_color_manual(values = c('in' = 'black', 'out' = 'red', 'adjacent' = 'orange')) +
         scale_fill_manual(values = c('tree' = 'lightgreen', 'jalon' = 'white')) +
         ggtitle(unique(tmp$sousplot)) +
         theme_classic() +
         theme(
            panel.grid.major = element_line(colour = "black"),
            panel.grid.minor  = element_line(colour = "white", linetype = "dotdash"),
            panel.background = element_rect(fill = "black")) +
         ggtitle(paste0('REPROJECTION FOR THE SUBPLOT :   ', unique(tmp$sousplot)))

      png(paste0(directory,'/',plot_name,"/20_20_PROJECTION/",'20_20_PROJECTION_', str_remove(unique(tmp$file),'.csv'),'.png'), width = 600)
      print(my_plot)
      dev.off()
   }
} # REPROJECT TREES TO 20 20 coordinates

dir.create(paste0(directory,'/',plot_name,"/Field_scan_comparison"))

{
   field_xy <- Fieldplot_BDD_full$extract %>%
      filter((!(is.na(position_x)) & !(is.na(position_y)))) %>%
      select(ind_num_sous_plot, position_x, position_y) %>%
      rename(id = ind_num_sous_plot) %>%
      mutate(id = as.character(id))

   all <- all %>% left_join(field_xy)

   all_comparison <- all %>% filter((!(is.na(position_x)) & !(is.na(position_y))) | what == 'jalon') 

   for (i in 1:length(unique(all_comparison$sousplot))) {

      tmp <- all_comparison %>%
         group_by(id) %>%
         arrange(match(method, c("method1", "method2")), .by_group = TRUE, across(starts_with("method2"))) %>%
         slice(1) %>%
         ungroup() %>%
         filter (sousplot == unique(all_comparison$sousplot)[i])

      tmp_tree <- tmp %>% filter(what == 'tree')
      tmp_jalon <- tmp %>% filter(what == 'jalon')

      my_plot <- ggplot() +
         ggrepel::geom_label_repel(data = tmp_tree, aes(label = id,x=TRUE_X, y=TRUE_Y), size = 5, fill = 'blue', col = 'black') +
         ggrepel::geom_label_repel(data = tmp_tree, aes(label = id,x=position_x, y=position_y), size = 5, fill = 'red', col = 'black') +
         ggrepel::geom_label_repel(data = tmp_jalon, aes(label = id,x=TRUE_X, y=TRUE_Y), size = 5, fill = 'green', col = 'black') +
         theme_classic() +
         theme(
            panel.grid.major = element_line(colour = "black"),
            panel.grid.minor  = element_line(colour = "white", linetype = "dotdash"),
            panel.background = element_rect(fill = "black")) +
         ggtitle(
            paste0('COMPARISON FIELD AND SCAN :   ', unique(tmp$sousplot)),
            subtitle = 'green : jalon,  blue : LIDAR,   red : manual')

      png(paste0(directory,'/',plot_name,"/Field_scan_comparison/",'Field_scan_comparison_', str_remove(unique(tmp$file),'.csv'),'.png'), width = 600)
      print(my_plot)
      dev.off()
   }

   } # XY compairison with field measurment

library(xlsx)
{
   write.xlsx(all, file = paste0(directory,'/',plot_name,"/20_20_PROJECTION.xlsx"), append = FALSE)
} # Save 20_20 projection

{
   tmp <- all %>%
      filter((!(is.na(position_x)) & !(is.na(position_y)))) %>%
      filter(!is.nan(TRUE_X))
   x_field <- tmp %>% .[['position_x']] %>% as.numeric()
   y_field <- tmp %>% .[['position_y']] %>% as.numeric()
   x_LIDAR <- tmp %>% .[['TRUE_X']] %>% as.numeric()
   y_LIDAR <- tmp %>% .[['TRUE_Y']] %>% as.numeric()

   coord_field <- as.matrix(cbind(x_field,y_field))
   coord_LIDAR <- as.matrix(cbind(x_LIDAR,y_LIDAR))

   pt_field <-    st_cast(st_sfc(st_multipoint(coord_field)), "POINT")
   pt_LIDAR <- st_cast(st_sfc(st_multipoint(coord_LIDAR)), "POINT")

   distance <- st_distance(pt_field,pt_LIDAR, by_element = TRUE)

   # hist(distance, xlab = 'Distance (m)', main = plot_name)

   png(paste0(directory,'/',plot_name,"/distance_", plot_name,'.png'), width = 600)
   hist(distance, breaks = c(0:45), xlab = 'Distance (m)', main = paste(plot_name, '  |  n = ', length(distance)))
   dev.off()

   tmp_dist <- tmp %>% mutate(distance = distance) %>% select(file, id, distance)

   write.xlsx(tmp_dist, file = paste0(directory,'/',plot_name,"/distance.xlsx"), append = FALSE)


} # Compute distance between field and scan measures
```


```{r}
print(paste('plots have been saved in : '))
print(paste0(getwd(),'/',plot_name,"/xy_rawdata_method1/"))
print(paste0(getwd(),'/',plot_name,"/xy_reproject_method1/"))
print('                                    ')
print(paste('A report has been saved : ', paste0(getwd(), '/',plot_name,"/3_method1_report.txt")))



```

```{r}

library(readxl)

all <- read_xlsx(paste0(directory,'/',plot_name,"/20_20_PROJECTION.xlsx"))
all <- all %>% 
   filter(what == 'tree' & where == 'in') %>% 
   group_by(id) %>% 
   slice(1) %>% 
   mutate(plot = "mbalmayo001") %>% 
   select(id,TRUE_X,TRUE_Y) 

all <- all %>% mutate(plot_name = plot_name)
write.xlsx(tmp_dist, file = paste0(directory,'/',plot_name,"/distance.xlsx"), append = FALSE)

write.xlsx(all %>% ungroup(), file = paste0(directory,'/',plot_name,"/mbalmayo001_final.xlsx"), append = FALSE)
 ```

