---
title: "extract_xy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{extract_xy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```


```{r setup}
library(xyLIDAR)
library(tidyverse)
library(BIOMASS)
library(plotsdatabase)
library(data.table)
library(grDevices)
library(sf)
library(writexl)
library(mapview)

load(system.file(package="xyLIDAR", 'color_plot.RData') )
load("C:/Users/hugol/Desktop/export_plot.RData")
load("C:/Users/hugol/Desktop/xyLIDAR/inst/color_plot.RData")

```

Set the parameters 

```{r}
# Parameters ####

root_in = "D:/VIA/scan_IPhone/mbalmayo004" #path to the folder of the lidar scans files
plot_name = "mbalmayo004" # plotsdatabase format
method = 'local_file' # Choose 'plotdatabase' or 'local_file'
crs = 'EPSG:32632'
directory <- 'D:/VIA/scan_IPhone/outputs'
```


```{r}
dir.create(file.path(directory,plot_name))
```

Import all the the lidar data files within a list (list_of_Lidar). Each element correspond to one file with 3 variables (id, X_lidar, Y_lidar)

files_Summary resumes the informations included in each files. If the infomation 
are not well organized in the dataframe, it means that the file is not well named.

```{r}
summary <- files_summary(root_in, 
                         export = TRUE,
                         directory = directory,
                         plot_name = plot_name)

list_of_Lidar <- summary$list_of_Lidar
files_Summary <- summary$file_summary

```


Extract field data 

```{r}
Fieldplot_BDD_full <- import_plotdata(method,plot_name) # Using plotsdatabase
# Fieldplot_BDD_full <- mba_004 # Or do it manually if you have it 
```

Compute jalons coordinates 

```{r}
subplot <- extract_jalons(coordinates = Fieldplot_BDD_full$coordinates, type = 2)
```

Reproject trees using the jalons coordinates as reference 

```{r}

all = XY_computation( 
      Fieldplot_BDD_full,
      list_of_Lidar,
      subplot,
      plot_name,
      files_summary,
      directory)

```

Plot all trees

```{r}

sub_plot <- plot_alltrees(
   Fieldplot_BDD_full,
   all,
   subplot,
   crs,
   col)

```


```{r}
save_gpkg(all, crs, sub_plot)
```

```{r}
all_scans_report(Fieldplot_BDD_full, all)
```

```{r}
all = reproj_20_20(all, directory, plot_name) 

all_scans_report (Fieldplot_BDD_full, all, directory, plot_name)

write_xlsx(all, path = paste0(directory,'/',plot_name,"/20_20_PROJECTION.xlsx")) # Save 20_20 projection

```


```{r}

library(readxl)

alli <- read_xlsx(paste0(directory,'/',plot_name,"/20_20_PROJECTION.xlsx"))
alli <- alli %>% 
   filter(what == 'tree' & where == 'in') %>% 
   group_by(id) %>% 
   slice(1) %>% 
   mutate(plot = "mbalmayo001") %>% 
   select(id,TRUE_X,TRUE_Y) 

alli <- alli %>% mutate(plot_name = plot_name)

# write.xlsx(all %>% ungroup(), file = paste0(directory,'/',plot_name,"/mbalmayo001_final.xlsx"), append = FALSE)
 ```

